How do galaxies get assigned to halos? \\ Searching for the assembly bias in the SDSS galaxy clustering

### 2016-09-16 text/ms.tex
newest draft

%%\documentclass[12pt, preprint]{aastex}
\documentclass[14pt, preprint]{emulateapj}
\usepackage{microtype}
\usepackage[breaklinks,colorlinks, urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{mathtools}
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue}

\DeclareMathOperator*{\argmax}{arg\,max}

\newcommand{\todo}[1]{{\em \textcolor{red}{ #1}}}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\lang}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\vep}{\bm{\epsilon}}
\newcommand{\ep}{\epsilon}
\newcommand{\pars}{\vec{\theta}}
\newcommand{\dev}{\mathrm{d}}
\newcommand{\wpp}{w_{\rm p}}
\newcommand{\ngal}{n_{g}}
\newcommand{\gmf}{g(N)}
\newcommand{\mr}{M_{\rm r}}
\newcommand{\rpp}{r_{\rm p}}
\newcommand{\mzero}{\log M_{0}}
\newcommand{\mone}{\log M_{1}}
\newcommand{\mmin}{\log M_{\rm min}}
\newcommand{\sigmam}{\sigma_{\log \rm M}} 
\newcommand{\acen}{\mathcal{A}_{\rm cen}}
\newcommand{\asat}{\mathcal{A}_{\rm sat}}

@@

\title{Can we constrain galaxy assembly bias with group and clustering statistics of SDSS galaxies?}
\author{Mohammadjavad Vakili \altaffilmark{1,$\dagger$}, David W. Hogg, ChangHoon~Hahn, Unspecified contributors}
\altaffiltext{$\dagger$}{\url{mjvakili@nyu.edu}}
\altaffiltext{1}{Center for Cosmology and Particle Physics, 
                 Department of Physics, New York University, 4 Washington Pl., New York, NY} 
\begin{abstract}

The standard halo occupation modeling approach in large scale structure 
studies assumes that the halo mass alone suffices in characterizing the 
galaxy-halo connection. Galaxy assembly bias, the hypothesis that properties beyond 
halo mass influences the galaxy-halo connection, can potentially affect the 
statistical summaries of galaxy populations. The degree to which, assembly bias can 
can be constrained with the statistical distribution of galaxies remains largely 
untested. In this investigation, we make use of the recently proposed assembly bias decoration 
of the HOD modeling in order to find constraints on the strength of the assembly bias from the most up-to-date and publicly available measurements of the group multiplicity function and the projected two-point correlation function of the Volume-limited luminosity threshold sample of SDSS 
galaxies. We find that the available GMF measurements cannot constrain galaxy assembly bias in any of the Luminosity thresholds. On the other hand, we find non-zero constraints on the satellite galaxy assembly bias for the $M_{r}<-18$ and $M_{r}<-19$ samples, nonzero constraints on the central galaxy assembly bias for the $M_{r}<-20$ samples, and constraints consistent with zero for the assembly bias of galaxies in the $M_{r}<-21$ sample. Furthermore, we provide a model comparison between the standard HOD approach and the decorated HOD in modeling of the GMF and the projected 2PCF measurements.

\end{abstract}

\section{Introduction}
The assumption that galaxies reside dark matter halos 
is the central tenet of the modern cosmology and 
large-scale structure formation theories. 
One of the most powerful methods for describing 
the galaxy-halo connection is the halo occupation modeling. 
The underlying assumption of the traditional HOD framework is 
that the halo mass alone is sufficient in determining the
population of galaxies inside halos. That is, the statistical properties of 
galaxies is governed by the halo mass. Mathematically, this assumption can be 
written as $P(N_g|M_h,\{x\})=P(N_g|M_h)$ where $\{x\}$ is the set of all 
possible halo properties beyond mass $M_{h}$. HOD provides an analytical prescription for the

Despite this very simplifying assumption, the galaxy-halo connection models such as HOD and CLF have been widely and successfully used in modeling the measurements of a wide range of the statistical properties of galaxies such as . However, the assumption that the statistical 
distribution of galaxies can be determined by halo mass is not a derived characteristic of 
accurate and high resolution hydrodynamic simulations, but rather a theoretical assumption 
that needs to be scrutinized. In recent years, the idea that the

In this investigation, we explore the notion that how the dependence of 
galaxy-halo connection on halo properties beyond mass (e.g. concentration, age).
Throughout this paper, unless stated otherwise, 
all radii and densities are in comoving units. Standard flat $\Lambda$CDM is assumed, 
and all cosmological parameters are set to the Planck 2013 best-fit estimates.

\section{Method}
\subsection{Halo occupation modeling}
\subsubsection{standard HOD modeling}\label{subsubsec:hod}
For our standard HOD modeling, HOD parameterization of cite[zheng07] is assumed. According to this parameterization, the occupation of the central galaxies follows a nearest-integer distribution, 
and the occupation of the satellite galaxies follows a Poisson distribution. The expected number of centrals and satellites as a function of the host halo mass of $M_{\rm h}$ are given by the following equations 
\begin{eqnarray}
\langle N_{\rm cen}(M_{\rm h}) \rangle &=& \frac{1}{2}\Big[1+\Big(\frac{\log M_{\rm h} - \log M_{\rm min}}{\sigma_{\log \rm{M}}} \Big) \Big], \label{hod:central}\\ 
\langle N_{\rm sat}(M_{\rm h}) \rangle &=& \Big( \frac{M_{\rm h} - M_{\rm{0}}}{M_{\rm 1}} \Big)^{\alpha}. \label{hod:satellite}
\end{eqnarray}

For populating the halos with galaxies, we follow the procedure described in CITE[HEARIN ETAL], CITE[CHANG ETAL]. The central galaxies are assumed to be at the center of the host dark matter halos. We place the satellite galaxies are within the virial radius of the halo following an NFW profile [CITE NFW]. The concentration of the NFW profile is given by the empirical mass-concentration relation provided by [CITE MACCIO]. The velocities of the satellite galaxies are given by two components. The first component is the velocity of the host halo. The second component is the velocity of the satellite galaxy with respect to the host halo which is computed following the solution to the NFW profile Jeans equations [CITE MORE]. We refer the readers to [CITE HEARIN 15] for a more comprehensive and detailed discussion of the forward modeling of the galaxy mock catalogs.

\subsubsection{Assembly bias HOD modeling}\label{subsubsec:decorated}
Now let us provide a brief overview of HOD modeling decorated with $\mathtt{Heaviside}$ $\mathtt{Assemblybias}$ introduced in [CITE HEARIN2015]. At a fixed halo mass $M_{\rm h}$, halos are split into two populations: population of halos with the 0.5-percentile of highest concentration, and population of halos with 0.5 percentile of lowest concentration.We call the first population "type 1" halos, and the second population "type 2 halos. Following In the decorated HOD model, the expected number of central and satellite galaxies in the two populations is given by

\begin{eqnarray}
\langle N_{c,i} | M_{h},c\rangle &=& \langle N_{c} | M_{h}\rangle + \delta N _{c,i}, \; i=1,2 \label{eq:decoratedcentral} \\
\langle N_{s,i} | M_{h},c\rangle &=& \langle N_{s} | M_{h}\rangle + \delta N _{s,i}, \; i=1,2 \label{eq:decoratedsatellite}
\end{eqnarray}
where $\langle N_{c} | M_{h}\rangle$ and $\langle N_{s} | M_{h}\rangle$ are given by Eqs \ref{hod:central} and \ref{hod:satellite} respectively, and we have $\delta N_{s,0} + \delta N_{s,1} = 0$, and $\delta N_{c,0} + \delta N_{c,1} = 0$. These two conditions ensure the conservation of HOD. At a given host hale mass $M_{h}$, the central occupation of the the two populations follows a Nearest-integer distribution with the first moment given by \ref{eq:decoratedcentral}; and the central occupation of the the two populations follows a Poisson distribution with the first moment given by \ref{eq:decoratedsatellite}.

In this decorated occupation model, the allowable range for the quantities $\delta N_{c,1}$ and $\delta N_{s,1}$ is given by 
\begin{eqnarray}
-\langle N_{c} | M_{h}\rangle \leq \delta N_{c} \leq \langle N_{c} | M_{h}\rangle
 , \label{eq:cen-bounds} \\
-\langle N_{s} | M_{h}\rangle \leq \delta N_{s} \leq \langle N_{s} | M_{h}\rangle. \label{eq:sat-bounds}
\end{eqnarray}
Afterwards, CITE[HEARIN2015] defines the assembly bias parameter $\mathcal{A}$ in the following way:
\begin{eqnarray}
\delta N_{\alpha , 1}(M_{h}) &=& \mathcal{A_\alpha} \delta N_{\alpha , 1}^{\rm max}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} > 0,  \\
\delta N_{\alpha , 1}(M_{h}) &=& \mathcal{A_\alpha} \delta N_{\alpha , 1}^{\rm min}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} < 0,
\end{eqnarray}
where the subscript $\alpha = c , s$ for the centrals and satellites respectively, and $\delta N_{\alpha , 1}^{\rm max}(M_{h})$, $\delta N_{\alpha , 1}^{\rm min}(M_{h})$ are given by Eqs. \ref{eq:cen-bounds} and \ref{eq:sat-bounds}.

At a given halo mass $M_{h}$, once the first moments of occupation statistics for the $type$-1 and $type$-2 halos are determined, the same procedure described in \ref{subsubsec:hod} is done to populate the halos with mock galaxies.

@@

We focus on two sets of measurements made on the volume limited luminosity threshold main sample of galaxies in the SDSS spectroscopic survey. In this section, we briefly describe the measurements used in our study for finding constraints on the assembly bias parameters.

\subsection{group multiplicity function}

>>>>

\documentclass[12pt, preprint]{aastex}
\usepackage[breaklinks,colorlinks, 
urlcolor=blue,citecolor=blue,linkcolor=blue]{hyperref}
\usepackage{graphicx}	% For figures
\usepackage{natbib}	% For citep and citep
\usepackage{amsmath}	% for \iint
\usepackage{mathtools}
\usepackage{bm}
\usepackage[breaklinks]{hyperref}	% for blackboard bold numbers
\usepackage{hyperref}
\usepackage{algorithmic,algorithm}
\hypersetup{colorlinks}
\usepackage{color}
\usepackage{morefloats}
\definecolor{darkred}{rgb}{0.5,0,0}
\definecolor{darkgreen}{rgb}{0,0.5,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\hypersetup{ colorlinks,
linkcolor=darkblue,
filecolor=darkgreen,
urlcolor=darkred,
citecolor=darkblue}

\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\lang}{\langle}
\newcommand{\ra}{\rangle}
\newcommand{\vep}{\bm{\epsilon}}
\newcommand{\ep}{\epsilon}
\newcommand{\pars}{\vec{\theta}}
\newcommand{\dev}{\mathrm{d}}
\newcommand{\ngal}{n_{g}}
\newcommand{\gmf}{g(N)}
\newcommand{\mr}{M_{\rm r}}
\newcommand{\rpp}{r_{\rm p}}
\newcommand{\mzero}{\log M_{0}}
\newcommand{\mone}{\log M_{1}}
\newcommand{\mmin}{\log M_{\rm min}}
\newcommand{\sigmam}{\sigma_{\log \rm M}} 
\newcommand{\wpp}{w_{\rm p}}

\newcommand{\acen}{\mathcal{A}_{\rm cen}}
\newcommand{\asat}{\mathcal{A}_{\rm sat}}
%\newcommand{\comment}[1]{{\bf \textcolor{red}{ #1}}}
\newcommand{\todo}[1]{{\em \textcolor{red}{ #1}}}

%% preprint2 produces a double-column, single-spaced document:

%% \documentclass[preprint2]{aastex}

%% Sometimes a paper's abstract is too long to fit on the
%% title page in preprint2 mode. When that is the case,
%% use the longabstract style option.

%% \documentclass[preprint2,longabstract]{aastex}

%% If you want to create your own macros, you can do so
%% using \newcommand. Your macros should appear before
%% the \begin{document} command.
%%
%% If you are submitting to a journal that translates manuscripts
%% into SGML, you need to follow certain guidelines when preparing
%% your macros. See the AASTeX v5.x Author Guide
%% for information.

%% You can insert a short comment on the title page using the command below.

%\slugcomment{Not to appear in Nonlearned J., 45.}

%% If you wish, you may supply running head information, although
%% this information may be modified by the editorial offices.
%% The left head contains a list of authors,
%% usually a maximum of three (otherwise use et al.).  The right
%% head is a modified title of up to roughly 44 characters.
%% Running heads will not print in the manuscript style.

%\shorttitle{Collapsed Cores in Globular Clusters}
%\shortauthors{Djorgovski et al.}

%% This is the end of the preamble.  Indicate the beginning of the
%% paper itself with \begin{document}.

@@

%% LaTeX will automatically break titles if they run longer than
%% one line. However, you may use \\ to force a line break if
%% you desire.
\title{How do galaxies get assigned to halos? \\ Searching for the assembly bias in the SDSS galaxy clustering}

%%\title{Collapsed Cores in Globular Clusters, \\
%%    Gauge-Boson Couplings, and AAS\TeX\ Examples}

%% Use \author, \affil, and the \and command to format
%% author and affiliation information.
%% Note that \email has replaced the old \authoremail command
%% from AASTeX v4.0. You can use \email to mark an email address
%% anywhere in the paper, not just in the front matter.
%% As in the title, use \\ to force line breaks.

\author{Mohammadjavad Vakili\altaffilmark{1}}

\email{mjvakili@nyu.edu}

\altaffiltext{1}{Center for Cosmology and Particle Physics, New York University, 4 Washington Pl, New York, NY, 10003}

\begin{abstract}

Clustering of dark matter halos is shown to depend on halo properties beyond mass such as halo concentration. This phenomenon is referred to as halo assembly bias. Standard halo occupation modeling in large scale structure studies assumes that halo mass alone is sufficient in characterizing the connection between galaxies and halos. Modeling of galaxy clustering can face systematic effects if galaxies are correlated with other halo properties. Using the Small MultiDark-Planck high resolution $N$-body simulation and the measurements of the projected two-point correlation function and the number density of Sloan Digital Sky Survey (SDSS) DR7 main galaxy sample, we investigate the extent to which the dependence of halo occupation on halo concentration can be constrained, and to what extent this can improve our modeling of galaxy clustering.

Given the SDSS clustering data, our constraints on HOD with assembly bias, suggests that satellite population is not correlated with halo concentration at fixed halo mass. Furthermore, in terms of the number of centrals at fixed halo mass, our results favor lack of correlation with halo concentration in the most luminous samples ($M_{\rm r}<-21.5,-21$), modest levels of correlation for $M_{\rm r}<-20.5,-20, -19.5$ samples, lack of correlation for $M_{\rm r}<-19,-18.5$ samples, and anti-correlation for the faintest sample $M_{\rm r}<-18$.

We show that in comparison with abundance matching mock catalogs, our findings suggest more modest levels of assembly bias. Furthermore, by performing model comparison based on information criteria, we find that in most cases, the standard mass-only HOD model is still favored by the observations.

\end{abstract}
\keywords{Cosmology: large-scale-structure of the universe, galaxies: halos}

\section{Introduction}

Most theories of cosmology and large scale structure formation under consideration today rely on the central assumption that galaxies reside in dark matter halos. Detailed study of the galaxy-halo connection is therefore critical in order to constrain cosmological models (by modeling galaxy clustering at non-linear scales) as well as providing a window into galaxy formation physics. One of the most powerful methods for describing the galaxy-halo connection is the halo occupation 
distribution (HOD, see \citealt{seljak2000, berlind_weinberg2002, scoccimarro2001, zheng2005, zheng07, leauthaud12, tinker2013, decorated, 2016arXiv160701782H}).

HOD is an empirical framework that provides an analytic prescription for the expected number of galaxies $N$ that reside in halos by specifying a probability distribution function $P(N|x)$ where $x$ is a property of the halo. Under standard HOD modeling, the assumption is made that the halo mass $M$ alone is sufficient in determining the galaxy population of a halo. That is, the statistical properties of galaxies is governed by the halo mass. Mathematically, this assumption can be written as $P(N_g|M,\{x\})=P(N_g|M)$ where $\{x\}$ is the set of all possible halo properties beyond halo mass $M$.

Despite this very simplifying assumption, the models of galaxy-halo connection based on HOD modeling have been used in fitting the measurements of a wide range of statistics such as the projected two-point correlation function of galaxies, small scale redshift space distortion, three-point function, and galaxy-galaxy lensing with remarkable success (e.g. \citealt{zheng07,tinker_rsd2007,zehavi2011,leauthaud12,parejko2013,coupon2015,hod-3pcf,guo2015,miyatake15,zu2015,hod_vs_sham}). 
HOD has been used in constraining the cosmological parameters through modeling the galaxy two-point correlation function (hereafter 2PCF) (\citealt{abazajian2005}), combination of 2PCF with mass-to-light ratio of galaxies (\citealt{tinker05}), redshift space distortions (\citealt{tinker_rsd2007}), mass-to-number ratio of galaxy clusters (\citealt{tinker2012}) galaxy-galaxy lensing (\citealt{vdb03,cacciato13,more13,vdb13}) in the main sample of galaxies
of the Sloan Digital Sky Survey (hereafter SDSS, \citealt{york2000}), and also the combination of galaxy clustering and galaxy-galaxy lensing (\citealt{more15}) in SDSS III Baryon Oscillation Spectroscopic Survey (BOSS, \citealt{boss}). Furthermore, HOD is implemented in producing mock galaxy catalogs in the BOSS survey (\citealt{manera2013,white2014}). It has also been used in galaxy evolution studies (\citealt{conroy09,leauthaud12,behroozi13,hudson2015,zu2015,zu2016}).

The simplifying assumption of standard HODs  
does not sufficiently model the complex small scale clustering and galaxy
formation. Numerous $N$-body simulations that examine the clustering of 
dark matter halos argue halo clustering is correlated with the formation 
history of halos. That is, at a fixed halo mass, the halo bias is correlated 
with properties of halos beyond mass such as the concentration, formation time, or etc. 
This phenomenon is known as the halo assembly bias (see \citealt{sheth2004,gao2005, harker2006, weschler2006, gao2007,croton2007,wang2007,angulo2008,dalal2008,li2008,sunayama2016}). 
There is even support for halo assembly bias in observations (\citealt{miyatake2016,more2016}).

Furthermore, the properties of galaxies within the halo, may also depend on the halo's formation history. Then, we may expect the occupation statistics of galaxies to be tied to the halo properties beyond mass such as the concentration of halos, environment, and etc. There have been many attempts in the literature in examining the dependence of the galaxy properties on larger scale environment of the halos. But the results are mixed and there is very little consensus over this matter. \citealt{tinker_void_2009} shows that the properties of the galaxies that reside in voids can be explained by the halo mass in which they live, and their properties are independent of their large scale environment of the halos. \citealt{tinker_density_hod} proposes and extension the standard HOD model $P(N|M)$ such that the number of galaxies residing in a halo not only depends on the mass of the halo, but also on the large scale density contrast $P(N|M,\delta)$. Based on modeling the clustering and void statistics of the SDSS galaxies, \citealt{tinker_vpf_2008} shows that the dependence of the expected number of central galaxies on large scale density is not very strong.    
By randomly shuffling the galaxies among host halos of similar mass in the Millenium simulation, \citealt{croton2007} shows that assembly bias significantly impacts the galaxy two-point correlation functions. They also show that the effect is different for the faint and the bright samples. By applying a group finder to SDSS DR7 galaxies, \citealt{tinker2011} shows that the quenched fraction of galaxies is independent of the large-scale environment.

Based on a group catalog of SDSS galaxies, \citealt{blanton_group_2007} finds that shuffling of galaxy colors among groups of similar mass has very little effect (only a slight change in small scales: less than $\sim 300 \rm{kpc}$) on the clustering of galaxies. Based on an alternative group catalog, \citealt{wang2013} show that clustering of the central galaxies does not depend only on the halo mass, but also on additional parameters such as the specific star formation rate.

Another family of empirical galaxy-halo connection models is \emph{Abundance} \emph{Matching}. In the Abundance Matching style models, galaxies are assumed to live in halos/subhalos and are assigned luminosities or stellar masses by assuming a one-to-one mapping that preserves the rank ordering between its host halo/subhalo property and its luminosity/stellar mass
(\citealt{kravtsov2004,vale2004,tasitsiomi2004,conroy2009,guo2010,wetzel2010,Neisten2011,watson2012,rodriguez2012,kravstov2013,mao2015,chavez2016}). 
In practice, the standard abundance matching models use host halo properties 
that trace the depth of the gravitational well of the halo, such as 
the maximum circular velocity the host halo had in its history, $V_\mathrm{max}$, 
and assume a scatter in this property's one-to-one mapping with 
stellar mass/luminosity. With this, 
they have been successfully used in modeling a wide range of the 
statistical properties of galaxies such as two-point correlation function (\citealt{reddick2013,lehman2015,hod_vs_sham}) as well as the group statistics of galaxies (\citealt{sham_gmf}).

It has been shown that the abundance matching mock catalogs that use $V_{\rm max}$ (see \citealt{hw2013,arz2014}), or the ones that use some combination of $V_{\rm max}$ and host halo/subhalo virial mass $M_{\rm vir} $(see \citealt{lehman2015}) exhibit significant levels of assembly bias. This has been demonstrated by randomizing the galaxies among host halos of in bins of halo mass, such that the HOD remains constant, and comparing the difference in the 2PCF of the randomized catalog and the original mock catalog.

Abundance Matching models, can be extended to account for even higher orders of galaxy 
assembly bias. In so called \emph{Age Matching} models, galaxy color is abundance-matched
to subhalo formation time so that older (redder) galaxies preferentially reside in older halos (\citealt{hw2013}). \citealt{hearin2014} demonstrates that such an approach can reproduce a range of color and luminosity-dependent statistics of SDSS galaxies such as the 2PCF and galaxy-galaxy lensing. Based on the projected 2PCF measurements of (\citealt{hw2013}) galaxy catalogs, \citealt{arz2014} showed that although the 2PCF measurements of these catalogs can be fit with the standard HOD modeling, the inferred HOD parameters cannot recover the underlying true HOD of these catalogs. That is, even in the presence of assembly bias in a galaxy sample, one can fit the clustering of this sample with HOD, but that does not guarantee recovery of the true HOD parameters. \citealt{mcewen2016} demonstrated that even with a galaxy catalog that has a significant level of assembly bias such as (\citealt{hw2013}) catalog, it is possible to extract the matter power spectrum more accurately if an environment-dependent HOD ($P(N|M,\delta)$) is used.

In this work, we aim to investigate the dependence of halo occupation on halo concentration, and how this dependence can be constrained in the low-redshift universe with the measurements of the 2PCF of galaxies with a wide range of luminosities in the SDSS DR7 main galaxy sample. 
In order to achieve this goal, we need to adopt a HOD model that takes into account 
a dependence on halo properties beyond mass. There are a number of frameworks in the literature. \citealt{edHOD-tinker,edHOD-gillmartin,edHOD-weinberg} have proposed environment-dependent HOD modeling methods that take into account the large-scale density contrast. The model that we use in this investigation is the decorated HOD modeling (\citealt{decorated}). In the decorated HOD framework, at a fixed halo mass, halos are populated with galaxies according to the standard HOD modeling. Then, a secondary halo property is selected. At a fixed halo mass, halos are split into two populations: halos with the highest and halos with the lowest values of the secondary property. The two population receive enhancements (decrements) in their number of galaxies according to a parameter which is the strength of the assembly bias parameter. Therefore, the assembly bias parameter will not be degenerate with the rest of the HOD parameters.

The advantage of this framework is that the more complex HOD model is identical to the underlying \emph{mass-only} HOD model in every respect, except that at a fixed halo mass, halos receive enhancement (decrements) in the number of galaxies they host according to the value of their secondary property. In order to constrain assembly bias along with the rest of the HOD parameters, we make use of the publicly available measurements of the projected 2PCF and number density measurements made by   
(\citealt{guo2015}). These measurements made use of the NYU value-added galaxy catalog (\citealt{Blanton2005}).

Furthermore, we discuss how and in which galaxy brightness limits, employing a more complex HOD model that takes assembly bias into account can improve our modeling of galaxy clustering. We discuss the physical mechanisms that enables the more concentration-dependent HOD model to improve fitting of the galaxy 2PCF. Then, we make qualitative comparison between the levels of assembly bias in the mock catalogs created from our constrained on the decorated HOD model, and the \citealt{hw2013} mock catalogs created using the abundance matching technique that uses maximum circular velocity $V_{\rm max}$ as subhalo property.

In order to investigate whether more complexity of the decorated HOD model is demanded by the galaxy clustering data, we perform model comparison between the standard HOD model and the HOD model with assembly bias. Moreover,  We discuss the effect of our choice of $N$-body simulation on our constraints, and previous works in the literature based on smaller $N$-body simulations (\citealt{zentner2016}).

This paper is structured as follows: In Section \ref{sec:method} we discuss the $N$-body simulation, the two halo occupation modeling methods, and the details of the computation of model observables used in this investigation. Then in Section \ref{sec:data}, we discuss the data used in this study. 
In Section \ref{sec:analysis} we discuss the details of our inference analysis as well as the results. This includes description of the details of our inference setup. In Section \ref{sec:result} we discuss the constraints and their implications. This includes presentation of the constraints on the parameters of the two models, interpretation of the predictions of our constraints and their possible physical ramifications, assessment of the levels of assembly bias as predicted by our model constraints and its comparison with abundance matching mock catalogs, and finally model comparison. Finally, we discuss and conclude in Section \ref{sec:summary}.
Throughout this paper, unless stated otherwise, 
all radii and densities are stated in comoving units. 
Standard flat $\Lambda$CDM is assumed, and all cosmological 
parameters are set to the Planck 2015 best-fit estimates.

\section{Method}\label{sec:method}

In this section, we discuss the ingredients of our modeling one-by-one. First, we discuss the simulation used in this study. Afterwards, we talk about the forward modeling of galaxy catalogs in the standard HOD modeling framework as well as the decorated HOD framework. Then, we provide an overview of the two summary statistics of the galaxy catalogs that we used in our inference.

For the simulations used in this work, we make use of the Rockstar (\citealt{rockstar}) halo catalogs in the $z=0$ snapshot of the Small MultiDark of Planck cosmology (referred to as $\mathtt{SMDP}$) (\citealt{smallmultidark}). This high resolution $N$-body simulation (publicly available at \url{https://www.cosmosim.org}) was carried out using the GADGET-2 code 
(see \citealt{smallmultidark} and references therein) code, following the Planck $\Lambda$CDM cosmological parameters 
$\Omega_{\rm m} = 0.307$, $\Omega_{\rm b} = 0.048$, $\Omega_{\Lambda} = 0.693$, $\sigma_{8} = 0.823$, $n_{\rm s}=0.96$, 
$h=0.678$. The Box size for this $N$-body simulation is 0.4 $h^{-1} \rm{Gpc}$, the number of simulation particles is 3840$^3$, the mass per simulation particle $m_{\rm p}$ is $9.6 \times 10^{7} \; h^{-1} M_{\odot}$, and the gravitational softening length $\epsilon$ is 1.5 $h^{-1} \rm{kpc}$.

In the $\mathtt{SMDP}$ simulation, as discussed in \citealt{halodemographic}, the Rockstar algorithm can reliably resolve halos with $\geq 100$ particles which corresponds to $M_{\rm vir} \geq 9.6 \times 10^{9} \; h^{-1}M_{\odot}$. The advantage of using the $\mathtt{SMDP}$ simulation is that it satisfies both the size and resolution requirements of studying the galaxy-halo connection. Fainter galaxy samples reside in lower mass halos, and the more luminous galaxy samples considered in this investigation occupy higher comoving volumes. Furthermore, since we are studying the higher order halo occupation statistics, the concentration-dependence in particular, it is important to use a simulation that can resolve the internal structure of halos. In the context of Subhalo-Abundance matching modeling---whose accuracy hinges upon subhalo completeness especially in the low mass limit---of the galaxy-halo connection in the SDSS data, $\mathtt{SMDP}$ has been previously used to model the faint galaxy samples with absolute luminosity in the r-band $M_{\rm r}$ less -18, -18.5, and -19 (see \citealt{hod_vs_sham}). The added advantage of using the $\mathtt{SMDP}$ simulation over some industrial simulation boxes commonly used in the literature such as $\mathtt{Bolshoi}$ (\citealt{Klypin2011,smallmultidark}) simulation is its comoving volume which makes it more suitable for performing inference with $L_{\star}$ (corresponding to $M_{\rm r} \sim -20.44$, see \citealt{blanton2003}) and more luminous than $L_{\star}$ galaxy samples.

\subsection{Halo occupation modeling}
\subsubsection{standard model without assembly bias}\label{subsubsec:hod}

For our standard HOD model, we assume the HOD parameterization from \citealt{zheng07}. 
According to this model, a dark matter halo can host a central galaxy and some number of satellite galaxies. The occupation of the central galaxies follows a nearest-integer distribution, 
and the occupation of the satellite galaxies follows a Poisson distribution. The expected number of centrals and satellites as a function of the host halo mass of $M_{\rm h}$ are given by the following equations 
\begin{eqnarray}
\langle N_{\rm cen}(M_{\rm h}) \rangle &=& \frac{1}{2}\Big[1+\Big(\frac{\log M_{\rm h} - \log M_{\rm min}}{\sigma_{\log \rm{M}}} \Big) \Big], \label{hod:central}\\ 
\langle N_{\rm sat}(M_{\rm h}) \rangle &=& \Big( \frac{M_{\rm h} - M_{\rm{0}}}{M_{\rm 1}} \Big)^{\alpha}. \label{hod:satellite}
\end{eqnarray}

For populating the halos with galaxies, we follow the procedure described in \citealt{2016arXiv160701782H,decorated}. The central galaxies are assumed to be at the center of the host dark matter halos. We assume that the central galaxies are at rest with respect to the bulk motion of the halos and their velocities are given by the velocity of the center of mass of their host halo. Note that this assumption is shown to be violated in brighter than $L_{\star}$ galaxies (see \citealt{guo2015}). But since we are not considering the redshift space 2PCF multipoles in our study, we do not expect this velocity bias to impact our inference. We place the satellite galaxies within the virial radius of the halo following a Navarro-Frenk-White profile (hereafter NFW; \citealt{nfw}). This approach is different from other simulation-based halo occupation modeling techniques (see \citealt{hod_vs_sham,zheng_guo}) in that the positions of the satellites are not assigned to the dark matter particles in the $N$-body simulation. 
The concentration of the NFW profile is given by the empirical mass-concentration relation provided by (\citealt{nfw_c(M)}). The velocities of the satellite galaxies are given by two components. The first component is the velocity of the host halo. The second component is the velocity of the satellite galaxy with respect to the host halo which is computed following the solution to the NFW profile Jeans equations (\citealt{more2010}). We refer the readers to \citealt{decorated} for a more comprehensive and detailed discussion of the forward modeling of the galaxy mock catalogs.

\subsubsection{Model with Assembly bias}\label{subsubsec:decorated}
Now let us provide a brief overview of HOD modeling with $\mathtt{Heaviside}$ $\mathtt{Assemblybias}$ (known as the decorated HOD) introduced in \citealt{decorated}. At a fixed halo mass $M_{h}$, halos are split into two populations: population of halos with the 0.5-percentile of highest concentration, and population of halos with 0.5 percentile of lowest concentration. For simplicity, we call the first population ``type-1'' halos, and the second population ``type-2'' halos. In the decorated HOD model, the expected number of central and satellite galaxies in the two populations are given by

\begin{eqnarray}
\langle N_{c,i} | M_{h},c\rangle &=& \langle N_{c} | M_{h}\rangle + \delta N _{c,i}, \; i=1,2 \label{eq:decoratedcentral} \\
\langle N_{s,i} | M_{h},c\rangle &=& \langle N_{s} | M_{h}\rangle + \delta N _{s,i}, \; i=1,2 \label{eq:decoratedsatellite}
\end{eqnarray}
where $\langle N_{c} | M_{h}\rangle$ and $\langle N_{s} | M_{h}\rangle$ are given by Eqs \ref{hod:central} and \ref{hod:satellite} respectively, and we have $\delta N_{s,1} + \delta N_{s,2} = 0$, and $\delta N_{c,1} + \delta N_{c,2} = 0$. These two conditions ensure the conservation of HOD. At a given host halo mass $M_{h}$, the central occupation of the the two populations follows a nearest-integer distribution with the first moment given by \ref{eq:decoratedcentral}; and the satellite occupation of the the two populations follows a Poisson distribution with the first moment given by \ref{eq:decoratedsatellite}.

In this occupation model, the allowable range that quantities $\delta N_{c,i}$ and $\delta N_{s,i}$ can take is given by 
\begin{eqnarray}
-\langle N_{c} | M_{h}\rangle \leq \delta N_{c,i} \leq \langle N_{c} | M_{h}\rangle
 , \label{eq:cen-bounds} \\
-\langle N_{s} | M_{h}\rangle \leq \delta N_{s,i} \leq \langle N_{s} | M_{h}\rangle. \label{eq:sat-bounds}
\end{eqnarray}
Afterwards, the assembly bias parameter $\mathcal{A}$ is defined in the following way:

\begin{eqnarray}
\delta N_{\alpha , 1}(M_{h}) &=& |\mathcal{A_\alpha}| \delta N_{\alpha , 1}^{\rm max}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} > 0,  \label{positive_A} \\
\delta N_{\alpha , 1}(M_{h}) &=& |\mathcal{A_\alpha}| \delta N_{\alpha , 1}^{\rm min}(M_{h}) \; \; \rm{if} \; \mathcal{A_\alpha} < 0, \label{negative_A}
\end{eqnarray}
where the subscript $\alpha = c , s$ for the centrals and satellites respectively, and $\delta N_{\alpha , 1}^{\rm max}(M_{h})$, $\delta N_{\alpha , 1}^{\rm min}(M_{h})$ are given by Eqs. \ref{eq:cen-bounds} and \ref{eq:sat-bounds}.

For a given $\mathcal{A}_{\alpha}$, once $\delta N_{\alpha,1}$ is computed using equation (\ref{positive_A})---if $\mathcal{A}_{\alpha}>0$---or equation (\ref{negative_A})---if $\mathcal{A}_{\alpha}<0$---, $\delta N_{\alpha,2} = 1 - \delta N_{\alpha,1}$ is computed. At a fixed halo mass $M_{h}$, once the first moments of occupation statistics for the $type$-1 and $type$-2 halos are determined, we perform the same procedure described in \ref{subsubsec:hod} to populate the halos with mock galaxies.

@@

We focus on the measurements made on the volume-limited luminosity-threshold main sample of galaxies in the SDSS spectroscopic survey. In this section, we briefly describe the measurements used in our study for finding constraints on the assembly bias as well as the HOD parameters.
